---
title: "level of masculinity"
author: "Nitay Alon"
date: "May 17, 2019"
output: html_document
---

In this notebook the goal is to compute the level of masculinity of an individual and a feature, that is, $p(i,j)$ denotes the probability of an individual $i$ demonstrate masculinity for feature $j$.

In the E-step of the EM algorithm we compute the above conditional probabilities. The first step is to extract these probabilities and create a new table of id and probabilities.

```{r load the data, include=FALSE}
local_server_prefix <- "~/Documents/Human_brain_research/DAPHNA_JOEL/"
source(paste0(local_server_prefix,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/analyze_full_brain_feature.R"))
source(paste0(local_server_prefix,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/method_helper.R"))
source(paste0(local_server_prefix,"BioBank_data_analysis/load_biobank_data.R"))

load("~/Documents/Human_brain_research/DAPHNA_JOEL/BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/Results/New_analysis_method_results_normalized.RData")
```

We need to keep the eid throughout the analysis process and export the table when the iterations are over.
```{r filter FDR significant features, include=FALSE}
sig_level <- 0.1
FDR_significant_features <- rownames(combined_df_long_sorted)[
    with(combined_df_long_sorted,(fdr_pv < sig_level))]  
```

extract feature data:
```{r extract FDR significant features data, include=FALSE}
significant_features <- results.list[FDR_significant_features[-193]]
men_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$men_responsebilities})
women_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$women_responsebilities})
```

join the tables:
```{r create men table, include=FALSE, warning=FALSE}
men_feature_data <- men_data[[1]]
for(i in 2:length(men_data))
{
  men_feature_data <- base::merge(men_feature_data, men_data[[i]], by.x = "eid", by.y = "eid")
}
names(men_feature_data) <- c("eid", FDR_significant_features[-193])
```

```{r create women table, include=FALSE, warning=FALSE}
women_feature_data <- women_data[[1]]
for(i in 2:length(women_data))
{
  women_feature_data <- base::merge(women_feature_data, women_data[[i]], by.x = "eid", by.y = "eid")
}
names(women_feature_data) <- c("eid", FDR_significant_features[-193])
```

## Correlation matrix

In order to determine if there's a clear pattern of masculinity among genders we compute the correlation matrix between each pair of features and then decompose the corrlation matrix to its eigenvalues to look for dependant.
```{r correlation matrix men}
library(corrplot)
men_cor <- cor(men_feature_data[,-c(1,94)])
upper_men_corr_matrix <- men_cor
upper_men_corr_matrix[lower.tri(men_cor,diag = T)] <- 0
which(upper_men_corr_matrix == max(upper_men_corr_matrix),arr.ind = T) 
colnames(upper_men_corr_matrix)[109]
rownames(upper_men_corr_matrix)[65]

men_correlation_matrix_eigen <- eigen(men_cor)
hist(men_correlation_matrix_eigen$values)
```

```{r correlation matrix women}
women_cor <- cor(women_feature_data[,-c(1,94)])
women_correlation_matrix_eigen <- eigen(women_cor)
hist(women_correlation_matrix_eigen$values)
```

We see the same pattern in the values of the eigen values among men and women.

