---
title: "level of masculinity"
author: "Nitay Alon"
date: "May 17, 2019"
output: html_document
---

In this notebook the goal is to compute the level of masculinity of an individual and a feature, that is, $p(i,j)$ denotes the probability of an individual $i$ demonstrate masculinity for feature $j$.

In the E-step of the EM algorithm we compute the above conditional probabilities. The first step is to extract these probabilities and create a new table of id and probabilities.

```{r load the data, include=FALSE}
# local_server_prefix <- "~/Documents/Human_brain_research/DAPHNA_JOEL/"
library(pbapply)
local_server_prefix <- "~/Human_brain_research/"
source(paste0(local_server_prefix,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/analyze_full_brain_feature.R"))
source(paste0(local_server_prefix,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/method_helper.R"))
source(paste0(local_server_prefix,"BioBank_data_analysis/load_biobank_data.R"))

load("~/Human_brain_research/BioBank_data_analysis/BioBank_data_analysis_total.RData")
load("~/Human_brain_research/BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/Results/combined_df_long_sorted.RData")
```

First select only FDR significant features from the antidiagonal: 
```{r filter FDR significant features, include=FALSE}
sig_level <- 0.01
FDR_significant_features <- rownames(combined_df_long_sorted)[
    with(combined_df_long_sorted,(fdr_pv <= sig_level & p < 0.5 & q > 0.5 |
                                  fdr_pv <= sig_level & p > 0.5 & q < 0.5))]  
FDR_significant_features <- na.omit(FDR_significant_features)
```

extract feature data:
```{r extract FDR significant features data, include=FALSE}
significant_features <- biobank_full_analysis_list[FDR_significant_features]

men_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$men_responsebilities})

women_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$women_responsebilities})
```

join the tables:
```{r create men table, include=FALSE, warning=FALSE}
men_feature_data <- men_data[[1]]
for(i in 2:length(men_data))
{
  men_feature_data <- base::merge(men_feature_data, men_data[[i]], by.x = "eid", by.y = "eid")
}
names(men_feature_data) <- c("eid", FDR_significant_features)
```

```{r create women table, include=FALSE, warning=FALSE}
women_feature_data <- women_data[[1]]
for(i in 2:length(women_data))
{
  women_feature_data <- base::merge(women_feature_data, women_data[[i]], by.x = "eid", by.y = "eid")
}
names(women_feature_data) <- c("eid", FDR_significant_features)
```

## Correlation matrix

In order to determine if there's a clear pattern of masculinity among genders we compute the correlation matrix between each pair of features and then decompose the corrlation matrix to its eigenvalues to look for dependant.
```{r correlation matrix men}
library(corrplot)
men_cor <- cor(men_feature_data[,-1])
hist(men_cor, breaks = 150, freq = F)
corrplot(men_cor, order = "FPC", addrect = 10,tl.cex = 0.6,tl.srt = 45)
```

```{r men feature histograms}
i <- sample(2:length(men_feature_data), 1)
hist(men_feature_data[,i],breaks = 150)
```


```{r men eigenvalues sum}
men_correlation_matrix_eigen <- eigen(men_cor)
ones <- rep(1,length(men_correlation_matrix_eigen$values))
men_correlation_matrix_eigen$vectors[,1]
hist(men_correlation_matrix_eigen$value, breaks = 50)
sum(log(men_correlation_matrix_eigen$values))
sum(log(ones))
```

```{r correlation matrix women}
women_cor <- cor(women_feature_data[,-1],use = "pairwise.complete.obs")
corrplot(women_cor, order = "FPC", addrect = 10,tl.cex = 0.6,tl.srt = 45)
women_correlation_matrix_eigen <- eigen(women_cor)
```

## Chi Square test for dependency:

```{r source chi square methods, include=FALSE}
source("../BioBank_data_analysis/measure_of_masculinity/chisq_between_features.R")
men_feature_data_disc <- pblapply(men_feature_data[,-1],
                                  ConvertProbabilityToBool) 

women_feature_data_disc <- pblapply(women_feature_data[,-1],
                                  ConvertProbabilityToBool) 
```


```{r test for dependency men, include=FALSE, warning=FALSE}
chisq_men <- matrix(NA,length(men_feature_data_disc),length(men_feature_data_disc))
not_valid_for_chisq_list_men <- c()

for(i in 1:(length(men_feature_data_disc)-1))
{
  for(j in (i+1):length(men_feature_data_disc))
  {
    data_for_chisq <- ComputeTwoByTwoTable(men_feature_data_disc[[i]],
                                       men_feature_data_disc[[j]])
    if(!is.null(data_for_chisq$valid) && data_for_chisq$valid)
    {
      chi_square_statistic <- (chisq.test(data_for_chisq$data))$statistic
      chi_square_sign <- ComputeChiSquareSign(data_for_chisq$data)
      chisq_men[i,j] <- sqrt(chi_square_statistic) * sign(chi_square_sign)
      next
    }
    not_valid_for_chisq_list_men <- rbind(not_valid_for_chisq_list_men,
                                       c(names(men_feature_data_disc)[i],
                                         names(men_feature_data_disc)[j]))
  }
}
```


```{r test for dependency women, include=FALSE, warning=FALSE}
chisq_women <- matrix(NA,length(men_feature_data_disc),length(men_feature_data_disc))
not_valid_for_chisq_list_women <- c()

for(i in 1:(length(women_feature_data_disc)-1))
{
  for(j in (i+1):length(women_feature_data_disc))
  {
    data_for_chisq <- ComputeTwoByTwoTable(women_feature_data_disc[[i]],
women_feature_data_disc[[j]])
    
    if(!is.null(data_for_chisq$valid) && data_for_chisq$valid)
    {
      chi_square_statistic <- (chisq.test(data_for_chisq$data))$statistic
      chi_square_sign <- ComputeChiSquareSign(data_for_chisq$data)
      chisq_women[i,j] <- sqrt(chi_square_statistic) * sign(chi_square_sign)
      next
    }
    not_valid_for_chisq_list_women <- rbind(not_valid_for_chisq_list_women,
                                      c(names(women_feature_data_disc)[i],
                                        names(women_feature_data_disc)[j]))
  }
}
```

Histograms of $\chi^2 * sign$:
```{r histograms of chi square times sign}
hist(chisq_men[!is.na(chisq_men)], breaks = 100,
     main = "Men", xlab = "Chi_square * sign")
hist(chisq_women[!is.na(chisq_women)], breaks = 100,
     main = "Women", xlab = "Chi_square * sign")
```

