---
title: "Analysis of main diagonal features"
author: "Nitay Alon"
date: "March 31, 2019"
output: html_document
---
In this report our goal is to examine the features found on the main diagonal of the $p-q$ plot.
```{r main diagonal features, include=FALSE}
library(ggplot2)
library(logcondens)
i <- sample(upper_main_diagonal_feature_names,1)
brain_feature <- results.list[[i]]
```

```{r histogram and CDF}
hist(brain_feature$feature_residuals$value, 
     breaks = 100,
     main = sprintf("Histogram of %s", upper_main_diagonal_feature_names[1]),
     xlab = "value", freq = FALSE)
plot(ecdf(brain_feature$feature_residuals$value), main = "CDF")
```

repeating the process for various features show similar picture - even after normalization there is a tail with equal proportion of men and women in these tails.

```{r histograms by gender}
ggplot(brain_feature$feature_residuals, aes(value)) + 
  facet_wrap(sex ~ .) +
  geom_histogram(bins = 150)
```

and computing non parametric log-likelihood by the formula:
\[
F_X(x) = \sum_{i=1}^{n}\mathbf{1}_{x < x_i}\frac{1}{n}
\]

```{r log likelihood ratio between men and women}
nonParametricLLRT(i)
```

Examination of the anti diagonal features:
```{r plot llrt anti diagonal}
brain_feature <- results.list[[upper_antidiagonal_feature_names[10]]]
plotGenderHistogram(brain_feature, upper_antidiagonal_feature_names[10])
qplot(brain_feature$feature_residuals$value, bins = 100)
nonParametricLLRT(upper_antidiagonal_feature_names[10], T)
```

