---
title: "level of masculinity"
author: "Nitay Alon"
date: "May 17, 2019"
output: html_document
---

In this notebook the goal is to compute the level of masculinity of an individual and a feature, that is, $p(i,j)$ denotes the probability of an individual $i$ demonstrate masculinity for feature $j$.

In the E-step of the EM algorithm we compute the above conditional probabilities. The first step is to extract these probabilities and create a new table of id and probabilities.

```{r load methods, include=FALSE}
# local_server_prefix <- "~/Documents/Human_brain_research/DAPHNA_JOEL/"
library(pbapply)
library(corrplot)
local_server_prefix <- "~/Human_brain_research/"
source(paste0(local_server_prefix,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/analyze_full_brain_feature.R"))
source(paste0(local_server_prefix,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/method_helper.R"))
# source(paste0(local_server_prefix,"BioBank_data_analysis/load_biobank_data.R"))
```

```{r load data, include=FALSE}
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/biobank_feature_residual_analysis.RData")
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/biobank_feature_standard_analysis.RData")
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/combined_df_long_sorted_residuals.RData")
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/combined_df_long_sorted_standardize.RData")
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/colomns_name.RData")
```

# Residual data

## Anti diagonal
First select only FDR significant features from the antidiagonal: 
```{r filter FDR significant features, include=FALSE}
sig_level <- 0.01
FDR_significant_features <- rownames(combined_df_long_sorted_residuals)[
    with(combined_df_long_sorted_residuals,(fdr_pv <= sig_level & 
region == "Anti"))]  
FDR_significant_features <- na.omit(FDR_significant_features)
```

extract feature data:
```{r extract FDR significant features data, include=FALSE}
significant_features <- biobank_feature_residual_analysis[FDR_significant_features]

men_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$men_responsebilities})

women_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$women_responsebilities})
```

```{r partiton to features, include=FALSE}
FA_features <- c(grep("Mean FA", names(significant_features)),
                 grep("mean FA", names(significant_features)))
MD_features <- c(grep("Mean MD", names(significant_features)),
                 grep("mean MD", names(significant_features)))
Volume_features <- c(grep("Volume", names(significant_features)))
feature_order_for_corrplot <- c(FA_features, MD_features, Volume_features)
```

join the tables:
```{r create men table, include=FALSE, warning=FALSE}
men_feature_data <- men_data[[1]]
for(i in 2:length(men_data))
{
  men_feature_data <- base::merge(men_feature_data, men_data[[i]], by.x = "eid", by.y = "eid")
}
names(men_feature_data) <- c("eid", FDR_significant_features)
men_features_no_id <- men_feature_data[,-1]
```

```{r create women table, include=FALSE, warning=FALSE}
women_feature_data <- women_data[[1]]
for(i in 2:length(women_data))
{
  women_feature_data <- base::merge(women_feature_data, women_data[[i]], by.x = "eid", by.y = "eid")
}
names(women_feature_data) <- c("eid", FDR_significant_features)
women_features_no_id <- women_feature_data[,-1]
```

## Cohen D histogram

```{r cohen D for FDR significant features}
hist(sapply(significant_features, function(x){x$hypothesis_results$cohen_d_test$estimate}),xlab = "Cohen-D", main = "Histogram of Cohen D", sub = "Residual data")
```


## Correlation matrix
```{r correlation matrix by gender, include=FALSE}
men_cor <- cor(men_features_no_id[,feature_order_for_corrplot], use = "pairwise.complete.obs")
women_cor <- cor(women_features_no_id[,feature_order_for_corrplot],use = "pairwise.complete.obs")
combined_correaltion <- men_cor
combined_correaltion[lower.tri(combined_correaltion)] <- women_cor[lower.tri(women_cor)]
```


```{r corrplot by name both genders}
corrplot(combined_correaltion, order="alphabet", title="both genders",
         tl.cex = 0.1,tl.srt = 45)
```

```{r corrplot by familiy both genders}
corrplot(combined_correaltion[FA_features,FA_features], method="color", title="FA",
         tl.cex = 0.1,tl.srt = 45)  
corrplot(combined_correaltion[MD_features,MD_features], method="color", title="MD",
         tl.cex = 0.1,tl.srt = 45)  
corrplot(combined_correaltion[Volume_features,Volume_features], method="color", title="Volume",
         tl.cex = 0.1,tl.srt = 45)  
```

# Standardize data

## Anti diagonal
First select only FDR significant features from the antidiagonal: 
```{r filter FDR significant features 2, include=FALSE}
sig_level <- 0.01
FDR_significant_features <- rownames(combined_df_long_sorted_standardize)[
    with(combined_df_long_sorted_standardize,(fdr_pv <= sig_level))]  
FDR_significant_features <- na.omit(FDR_significant_features)
```

extract feature data:
```{r extract FDR significant features data 2, include=FALSE}
significant_features <- biobank_feature_standard_analysis[FDR_significant_features]

men_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$men_responsebilities})

women_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$women_responsebilities})
```

join the tables:
```{r create men table 2, include=FALSE, warning=FALSE}
men_feature_data <- men_data[[1]]
for(i in 2:length(men_data))
{
  men_feature_data <- base::merge(men_feature_data, men_data[[i]], by.x = "eid", by.y = "eid")
}
names(men_feature_data) <- c("eid", FDR_significant_features)
```

```{r create women table 2, include=FALSE, warning=FALSE}
women_feature_data <- women_data[[1]]
for(i in 2:length(women_data))
{
  women_feature_data <- base::merge(women_feature_data, women_data[[i]], by.x = "eid", by.y = "eid")
}
names(women_feature_data) <- c("eid", FDR_significant_features)
```

## Cohen D histogram

```{r cohen D for FDR significant features 2}
hist(sapply(significant_features, function(x){x$hypothesis_results$cohen_d_test$estimate}),xlab = "Cohen-D", main = "Histogram of Cohen D", sub = "Residual data")
```


## Correlation matrix
```{r correlation matrix by gender 2, include=FALSE}
men_cor <- cor(men_features_no_id[,feature_order_for_corrplot], use = "pairwise.complete.obs")
women_cor <- cor(women_features_no_id[,feature_order_for_corrplot],use = "pairwise.complete.obs")
combined_correaltion <- men_cor
combined_correaltion[lower.tri(combined_correaltion)] <- women_cor[lower.tri(women_cor)]
```

```{r corrplot by name both genders 2}
corrplot(combined_correaltion, order="alphabet", title="both genders",
         tl.cex = 0.1,tl.srt = 45)
```

```{r corrplot by familiy both genders 2}
corrplot(combined_correaltion[FA_features,FA_features], method="color", title="FA",
         tl.cex = 0.1,tl.srt = 45)  
corrplot(combined_correaltion[MD_features,MD_features], method="color", title="MD",
         tl.cex = 0.1,tl.srt = 45)  
corrplot(combined_correaltion[Volume_features,Volume_features], method="color", title="Volume",
         tl.cex = 0.1,tl.srt = 45)  
```
