---
title: "level of masculinity"
author: "Nitay Alon"
date: "May 17, 2019"
output: html_document
---

In this notebook the goal is to compute the level of masculinity of an individual and a feature, that is, $p(i,j)$ denotes the probability of an individual $i$ demonstrate masculinity for feature $j$.

In the E-step of the EM algorithm we compute the above conditional probabilities. The first step is to extract these probabilities and create a new table of id and probabilities.

```{r load methods, include=FALSE}
# local_server_prefix <- "~/Documents/Human_brain_research/DAPHNA_JOEL/"
library(pbapply)
library(corrplot)
local_server_prefix <- "~/Human_brain_research/"
source(paste0(local_server_prefix,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/analyze_full_brain_feature.R"))
source(paste0(local_server_prefix,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/method_helper.R"))
source(paste0(local_server_prefix,"BioBank_data_analysis/load_biobank_data.R"))
```

```{r load data, include=FALSE}
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/biobank_feature_residual_analysis.RData")
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/biobank_feature_standard_analysis.RData")
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/combined_df_long_sorted_residuals.RData")
load("~/Human_brain_research/BioBank_data_analysis/Final_data_for_analysis/combined_df_long_sorted_standardize.RData")
```

# Residual data

First select only FDR significant features from the antidiagonal: 
```{r filter FDR significant features, include=FALSE}
sig_level <- 0.01
FDR_significant_features <- rownames(combined_df_long_sorted_residuals)[
    with(combined_df_long_sorted,(fdr_pv <= sig_level & p < 0.5 & q > 0.5 |
                                  fdr_pv <= sig_level & p > 0.5 & q < 0.5))]  
FDR_significant_features <- na.omit(FDR_significant_features)
```

extract feature data:
```{r extract FDR significant features data, include=FALSE}
significant_features <- biobank_feature_residual_analysis[FDR_significant_features]

men_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$men_responsebilities})

women_data <- lapply(significant_features, function(x){x$hypothesis_results$pure_type_vs_mixed_gender_em_results$alternative_hypothesis$women_responsebilities})
```

join the tables:
```{r create men table, include=FALSE, warning=FALSE}
men_feature_data <- men_data[[1]]
for(i in 2:length(men_data))
{
  men_feature_data <- base::merge(men_feature_data, men_data[[i]], by.x = "eid", by.y = "eid")
}
names(men_feature_data) <- c("eid", FDR_significant_features)
```

```{r create women table, include=FALSE, warning=FALSE}
women_feature_data <- women_data[[1]]
for(i in 2:length(women_data))
{
  women_feature_data <- base::merge(women_feature_data, women_data[[i]], by.x = "eid", by.y = "eid")
}
names(women_feature_data) <- c("eid", FDR_significant_features)
```

## Cohen D histogram

```{r cohen D for FDR significant features}
hist(sapply(significant_features, function(x){x$hypothesis_results$cohen_d_test$estimate}),xlab = "Cohen-D", main = "Histogram of Cohen D", sub = "Residual data")
```


## Correlation matrix

In order to determine if there's a clear pattern of masculinity among genders we compute the correlation matrix between each pair of features and then decompose the corrlation matrix to its eigenvalues to look for dependant.
```{r correlation matrix by gender, include=FALSE}
which(names(men_feature_data) == "X25901.2.0")
men_cor <- cor(men_feature_data[,-c(1,64)], use = "pairwise.complete.obs")
women_cor <- cor(women_feature_data[,-c(1,64)],use = "pairwise.complete.obs")
corrplot(women_cor, order = "FPC", addrect = 10,tl.cex = 0.6,tl.srt = 45)
```

```{r plot correlation by gender}
corrplot(men_cor, order = "FPC", addrect = 10,tl.cex = 0.6,tl.srt = 45, title = "Men")
corrplot(women_cor, order = "FPC", addrect = 10,tl.cex = 0.6,tl.srt = 45, title = "Women")
```

```{r corrplot by name both genders}
combined_correaltion <- men_cor
# replace the values
combined_correaltion[lower.tri(combined_correaltion)] <- women_cor[lower.tri(women_cor)]

corrplot(combined_correaltion, order="alphabet", title="both genders")
```

# By quadrant analysis
In this subsection our goal is to establish 

<!-- ```{r men feature histograms} -->
<!-- i <- sample(2:length(men_feature_data), 1) -->
<!-- hist(men_feature_data[,i],breaks = 150) -->
<!-- ``` -->


<!-- ```{r men eigenvalues sum} -->
<!-- men_correlation_matrix_eigen <- eigen(men_cor) -->
<!-- ones <- rep(1,length(men_correlation_matrix_eigen$values)) -->
<!-- men_correlation_matrix_eigen$vectors[,1] -->
<!-- hist(men_correlation_matrix_eigen$value, breaks = 50) -->
<!-- sum(log(men_correlation_matrix_eigen$values)) -->
<!-- sum(log(ones)) -->
<!-- ``` -->

<!-- ```{r correlation matrix women} -->
<!-- corrplot(women_cor, order = "alphabet", addrect = 10,tl.cex = 0.6,tl.srt = 45) -->
<!-- ``` -->

<!-- ## Chi Square test for dependency: -->

<!-- ```{r source chi square methods, include=FALSE} -->
<!-- source("../BioBank_data_analysis/measure_of_masculinity/chisq_between_features.R") -->
<!-- men_feature_data_disc <- pblapply(men_feature_data[,-1], -->
<!--                                   ConvertProbabilityToBool)  -->

<!-- women_feature_data_disc <- pblapply(women_feature_data[,-1], -->
<!--                                   ConvertProbabilityToBool)  -->
<!-- ``` -->


<!-- ```{r test for dependency men, include=FALSE, warning=FALSE} -->
<!-- chisq_men <- matrix(NA,length(men_feature_data_disc),length(men_feature_data_disc)) -->
<!-- not_valid_for_chisq_list_men <- c() -->

<!-- for(i in 1:(length(men_feature_data_disc)-1)) -->
<!-- { -->
<!--   for(j in (i+1):length(men_feature_data_disc)) -->
<!--   { -->
<!--     data_for_chisq <- ComputeTwoByTwoTable(men_feature_data_disc[[i]], -->
<!--                                        men_feature_data_disc[[j]]) -->
<!--     if(!is.null(data_for_chisq$valid) && data_for_chisq$valid) -->
<!--     { -->
<!--       chi_square_statistic <- (chisq.test(data_for_chisq$data))$statistic -->
<!--       chi_square_sign <- ComputeChiSquareSign(data_for_chisq$data) -->
<!--       chisq_men[i,j] <- sqrt(chi_square_statistic) * sign(chi_square_sign) -->
<!--       next -->
<!--     } -->
<!--     not_valid_for_chisq_list_men <- rbind(not_valid_for_chisq_list_men, -->
<!--                                        c(names(men_feature_data_disc)[i], -->
<!--                                          names(men_feature_data_disc)[j])) -->
<!--   } -->
<!-- } -->
<!-- ``` -->


<!-- ```{r test for dependency women, include=FALSE, warning=FALSE} -->
<!-- chisq_women <- matrix(NA,length(men_feature_data_disc),length(men_feature_data_disc)) -->
<!-- not_valid_for_chisq_list_women <- c() -->

<!-- for(i in 1:(length(women_feature_data_disc)-1)) -->
<!-- { -->
<!--   for(j in (i+1):length(women_feature_data_disc)) -->
<!--   { -->
<!--     data_for_chisq <- ComputeTwoByTwoTable(women_feature_data_disc[[i]], -->
<!-- women_feature_data_disc[[j]]) -->

<!--     if(!is.null(data_for_chisq$valid) && data_for_chisq$valid) -->
<!--     { -->
<!--       chi_square_statistic <- (chisq.test(data_for_chisq$data))$statistic -->
<!--       chi_square_sign <- ComputeChiSquareSign(data_for_chisq$data) -->
<!--       chisq_women[i,j] <- sqrt(chi_square_statistic) * sign(chi_square_sign) -->
<!--       next -->
<!--     } -->
<!--     not_valid_for_chisq_list_women <- rbind(not_valid_for_chisq_list_women, -->
<!--                                       c(names(women_feature_data_disc)[i], -->
<!--                                         names(women_feature_data_disc)[j])) -->
<!--   } -->
<!-- } -->
<!-- ``` -->

<!-- Histograms of $\chi^2 * sign$: -->
<!-- ```{r histograms of chi square times sign} -->
<!-- hist(chisq_men[!is.na(chisq_men)], breaks = 100, -->
<!--      main = "Men", xlab = "Chi_square * sign") -->
<!-- hist(chisq_women[!is.na(chisq_women)], breaks = 100, -->
<!--      main = "Women", xlab = "Chi_square * sign") -->
<!-- ``` -->

<!-- ## Correlation matrix of features -->

<!-- Unlike the above analysis, in this section we compute the correlation matrix between the residuals and not the probabilities: -->
<!-- ```{r extract men residual for corrplot, include=FALSE, warning=FALSE} -->
<!-- men_residuals <- sapply(significant_features, function(x){x$feature_residuals$value[x$feature_residuals$sex == 1]}) -->

<!-- women_residuals <- sapply(significant_features, function(x){x$feature_residuals$value[x$feature_residuals$sex == 0]}) -->

<!-- men_residuals_matrix <- do.call(cbind, men_residuals) -->
<!-- names(men_residuals_matrix) = names(men_residuals) -->

<!-- women_residuals_matrix <- do.call(cbind, women_residuals) -->
<!-- names(women_residuals_matrix) = names(women_residuals) -->
<!-- ``` -->

<!-- ```{r computing correlation between residuals men, include=TRUE} -->
<!-- men_residuals_corr_matrix <- cor(men_residuals_matrix, -->
<!--                                  use = "pairwise.complete.obs") -->
<!-- corrplot(men_residuals_corr_matrix, order = "alphabet", addrect = 10,tl.cex = 0.6,tl.srt = 45) -->
<!-- corrplot(men_residuals_corr_matrix, order = "FPC", addrect = 10,tl.cex = 0.6,tl.srt = 45) -->
<!-- ``` -->

<!-- ```{r computing correlation between residuals women, include=TRUE} -->
<!-- women_residuals_corr_matrix <- cor(women_residuals_matrix, -->
<!--                                  use = "pairwise.complete.obs") -->
<!-- corrplot(women_residuals_corr_matrix, order = "alphabet", addrect = 10,tl.cex = 0.6,tl.srt = 45) -->
<!-- corrplot(women_residuals_corr_matrix, order = "FPC", addrect = 10,tl.cex = 0.6,tl.srt = 45) -->
<!-- ``` -->

