---
title: "Exploring the distribution of gender proportions in mixture model"
author: "Nitay Alon"
date: "2/28/2019"
output: html_document
---

In this report we explore the joint distribution (structure) of the mixture proportions of brain features volumes. In the first section a short explanation on the meaning of the proportions is given followed by a set of plots exploring the structure of these variables. 

# Gender preference proprtions
The mixture model assumes that the distribution of each brain feature measures is a mixture of 2 underline normal distributions (assuming unequal variance and mean). Since the data is scaled one of those distributions will have a mean over 0 and the other one below it. 

In turn each gender "selects" the proportion of high mean distribution and low mean distribution, denoted $p,q$ in this report. The goal of this report is to explore the distribution of these proportions.

# Joint distribution of proportions:
```{r fit quadrant, warning=FALSE,include=FALSE}
fitQuadrant <- function(p,q)
{
  p_tilde <- as.numeric(p) - 0.5
  q_tilde <- as.numeric(q) - 0.5
  if(sign(p_tilde) == sign(q_tilde))
  {
    return("Main")
  }
  return("Anti")
}
```

```{r sourcing methods, warning=FALSE,include=FALSE}
library(ggplot2)
library(dplyr)
library(Jmisc)
library(config)
main_folder_path <- switch (
    (Sys.info())[[4]],
    "nitayServer"="~/Documents/Human_brain_research/DAPHNA_JOEL/",
    "Nitay-R-Droplet"="~/HumanBrainResearch/"
  )
  load(paste0(main_folder_path,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/Results/New_analysis_method_results_normalized.RData"))
  source(paste0(main_folder_path,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/convert_em_probs_to_pq_space.R"))
  source(paste0(main_folder_path,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/merge_data_frames_by_column_names.R"))
  source(paste0(main_folder_path,"BioBank_data_analysis/New_Analysis_Method/Residuals_analysis_methods/convert_em_probs_to_pq_space.R"))
  sourceAll(paste0(main_folder_path,"BioBank_data_analysis/Visualization_methods"))
```

```{r prepare data for analysis, include = FALSE}
results_list_m_parameters <- convertEMProbabiltiesToPQSpace(results.list)
p_and_q <- sapply(results_list_m_parameters, function(x){c(x$p, x$q)})
```

### P-Value color schema pure vs mixture

First we add a coloring schema based on the *p-value* of the llrt, when the hypothesis is pure type vs mixture model. 
```{r pure type vs mixture model colouring schema, include=FALSE,warning=FALSE}
pvalue_per_feature <- pullPvaluePerHypothesis(results.list)
pvalue_per_feature_df <- t(pvalue_per_feature) %>% as.data.frame() 
p_and_q_df <- p_and_q %>% as.data.frame()
combined_df <- mergeDataFramesByColumnNames(pvalue_per_feature_df, p_and_q_df)
rownames(combined_df) <- c("p_values","p","q")
```

```{r pure type vs mixture model plotting with colors}
combined_df_long <- combined_df %>% t() %>% as.data.frame()
combined_df_long$bins_of_pv <- cut(combined_df_long$p_values, c(1e-9,1e-8,1e-3,1e-2,1))
ggplot(combined_df_long, aes(p,q, colour = bins_of_pv)) + 
  geom_point() + 
  ggtitle("Values of p and q", subtitle = "No correction")
```

### FDR

The following figure plots the FDR corrected $p-values$, point color indicate feature quadrant, horizontal lines indicate significance levels.
```{r fdr procedure, include = FALSE}
N <- nrow(combined_df_long)
combined_df_long_sorted <- combined_df_long[order(combined_df_long$p_values,decreasing = F),]
combined_df_long_sorted$i <- seq(1:N)
combined_df_long_sorted$bh_pvalue <- with(combined_df_long_sorted, p_values * i / N)
combined_df_long_sorted$region <- apply(combined_df_long_sorted, 1, function(x){fitQuadrant(x[2],x[3])})
```

```{r plot fdr}
ggplot(combined_df_long_sorted, aes(i,bh_pvalue,colour = region)) + 
  geom_point() +
  geom_hline(yintercept = 0.01) +
  geom_hline(yintercept = 0.05) +
  geom_hline(yintercept = 0.1) +
  ggtitle("FDR p-values")
  # geom_vline(xintercept = 0.5) +
  # geom_abline(slope = 1, intercept = 0)
```

The tables below are the number of significant features, per significance level:

```{r summarize of significant features}
table(combined_df_long_sorted$bh_pvalue < 0.01)
table(combined_df_long_sorted$bh_pvalue < 0.05)
table(combined_df_long_sorted$bh_pvalue < 0.1)
```

To compute the FDR $p-value$ the current p-values are transformed using the formula:
\[
  pv(i) = max_{j \leq i} pv \frac{n}{j} 
\]

```{r fdr pvalues, include = FALSE}
combined_df_long_sorted$fdr_pv <- 
  sapply(c(1:nrow(combined_df_long_sorted)), function(i){
    max(combined_df_long_sorted$bh_pvalue[1:i])
  })
  
combined_df_long_sorted$bins_for_fdr <- cut(combined_df_long_sorted$fdr_pv, c(0,1e-2,5e-2,1e-1,1))
```

```{r plot fdr pvalues, warnings = FALSE}
ggplot(combined_df_long_sorted, aes(i,fdr_pv, colour = bins_for_fdr)) + 
  geom_point() +
  ggtitle("FDR pv", subtitle = "color indicate sig level") 
```
```{r fdr p value by quadrant, warnings = FALSE}
ggplot(combined_df_long_sorted, aes(p,q, colour = bins_for_fdr)) + 
  geom_point() +
  geom_vline(xintercept = 0.5) +
  geom_hline(yintercept = 0.5) +
  ggtitle("Values of p and q", subtitle = "color indicate FDR pv") + 
  geom_abline(slope = 1, intercept = 0)
```

The above measles like plot shows the distribution of the FDR p-values. It is visible that the II and IV quadrants are heavily dense with significant features. Some features in the I quadrant and the III quadrants are also significant. 

# Post hoc analysis

In this section our goal is to find patterns in the distribution structure of the FDR significance features. Assuming that the underline model is a mixture of Gaussians with different means and variance, these feature should present a distinct pattern. Using visual inspection of the underline distribution it is possible to uncover characteristics of the features. The analysis begins with the anti-diagonal ($sign(p-0.5) \neq sign(q-0.5)$) followed by the same method for the main diagonal.

## antidiagonal

### Upper
```{r high q low p}
upper_antidiagonal_feature_names <- 
  rownames(combined_df_long_sorted)[
    with(combined_df_long_sorted,(fdr_pv < 5e-2 & p < 0.5 & q > 0.5))]  
save(upper_antidiagonal_feature_names, file = "upper_diagonal_feature_names.RData")    
upper_antidiagonal_features <- 
  results.list[names(results.list) %in% upper_antidiagonal_feature_names]
```

```{r plotting upper anti diagonal feature}
i <- sample(1:length(upper_antidiagonal_feature_names), 1)
plotGenderHistogram(upper_antidiagonal_features[[i]], upper_antidiagonal_feature_names[i])
```

A general pattern in this qudrant is best represented by the following plot:
```{r representitive plot}
feature_name = "X25805.2.0"
plotGenderHistogram(upper_antidiagonal_features[[feature_name]], feature_name)
```

### Lower 
```{r low q high p}
lower_antidiagonal_feature_names <- 
  rownames(combined_df_long_sorted)[
    with(combined_df_long_sorted,(fdr_pv < 1e-3 & p > 0.5 & q < 0.5))]  
save(lower_antidiagonal_feature_names, file = "lower_diagonal_feature_names.RData")  
lower_antidiagonal_features <- 
  results.list[names(results.list) %in% lower_antidiagonal_feature_names]
```

```{r plotting at random}
i <- sample(1:length(lower_antidiagonal_feature_names), 1)
plotGenderHistogram(lower_antidiagonal_features[[i]], lower_antidiagonal_feature_names[i])
```

```{r high mixture for men wemon neutral}
feature_name = "X25120.2.0"
plotGenderHistogram(lower_antidiagonal_features[[feature_name]], feature_name)
```


```{r another example of sex mixture model}
feature_name = "X25535.2.0"
plotGenderHistogram(lower_antidiagonal_features[[feature_name]], feature_name)
```


<!-- ## Main diagonal -->
<!-- Analyzing the features based on the *p-values*: -->
<!-- ```{r upper main diagonal features} -->
<!-- upper_main_diagonal_feature_names <-  -->
<!--   rownames(combined_df_long_sorted)[ -->
<!--     with(combined_df_long_sorted,(p_values < 1e-3 & p > 0.5 & q > 0.5))]   -->
<!-- setwd("~/Documents/Human_brain_research/DAPHNA_JOEL/BioBank_data_analysis/Log_likelihood_ratio_computation") -->
<!-- save(upper_main_diagonal_feature_names, file = "upper_diagonal_feature_names.RData") -->
<!-- upper_main_diagonal_features <-  -->
<!--   results.list[names(results.list) %in% upper_main_diagonal_feature_names] -->
<!-- ``` -->

<!-- Applying the same plotting schema as in the shiny app: -->
<!-- ```{r plotting upper diagonal feature} -->
<!-- i <- 3 -->
<!-- plotGenderHistogram(upper_main_diagonal_features[[i]], upper_main_diagonal_feature_names[i]) -->
<!-- ``` -->

<!-- it is clear that in the case of the upper diagonal the mixture model yields higher likelihood ratio only due to right tailed observations. -->

<!-- ### Non parametric LLRT -->
<!-- ```{r non parametric llrt of main diagonal} -->
<!-- library(logcondens) -->
<!-- men_density <- logConDens() -->
<!-- ``` -->





<!-- ```{r find interesting regions} -->
<!-- findInterestingRegion(lower_antidiagonal_features) -->
<!-- ``` -->

